---
title: "top_levels_function"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

## Learning how to use the top_levels() function in the janitor package  
top_levels() is helpful to look at a large dataset and see which values or variables are most frequent. You can also ask R to specify how many "top levels" you want to see.  
For example, let's say we have a dataset with:  
-  1,1,1,2,3,3,3,3,3,4,5

When you use the top_level() function, R would tell you that you the value 3 is the most common within the dataset and if you wanted to specify what the top 2 most common values were, R would include 3 and 1.  

Pretty cool! 

Let's try with some actual code and data now :) 

```{r}
library(palmerpenguins)
library(janitor)

#top_levels(penguins$species, 2)

??top_levels
```

## Some notes to consider:  
-  you can only use top_levels() with factors! It will not work as a numeric value 
-  make sure to install the janitor package and then load the library before trying to use top_levels() function
-  the category must also have at least 3 levels 
-  this warning message means there is an overlap in the top and bottom groups - use a different dataset for example 









