---
title: "Top Levels Function, top_levels()"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: 
  html_document: 
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.path= "../Outputs/")
```

### Introduction:  
-  In this section, we are going to learn how to use the top_levels() function in the janitor package. Have you ever had a large dataset and wanted to be able to look at which values or variable are most frequent? 
-  Well, using **top_levels()** can help with that! 

You can also ask R to specify how many "top levels" you want to see.  

-  For example, let's say we have a dataset with: 1,1,1,2,3,3,3,3,3,4,5

When you use the top_level() function, R would return and tell you that the value "3" is the most common within the dataset and if you wanted to specify what the _top 2_ most common values were, R would output 3 and 1.  

Pretty cool! 

#### Let's try with some actual code and data now :) 

```{r, warning=FALSE}
## Load the libraries that are necessary to demonstrate this function 
library(palmerpenguins)
library(janitor)

#top_levels(penguins$species)
## this is an important example to use because it shows that there is an overlap in the top and bottom groups. 
## So maybe we should use a different dataset with more levels...

## If you are ever unsure or forget how the package or function works, remember you can use question marks to get more information about the uses of the function. 
??top_levels
```

### Let's try with a larger dataset:
```{r}
## Read in the dataset for this example
library(here)
library(tidyverse)
drugs <- read_csv(here("Hannah", "Data", "drugs.csv"))
MCR <- read_csv(here("Hannah", "Data", "MCR_JCU_PLoS_DataArchive_20150121.csv"))

## A good way to check to see if the variable you are hoping to filter has enough categories is by using the distinct() function 
drugs %>% 
  distinct(category)
# so category of drug would not work here because there are only 2 

# but therapeutic area may be a good example to work with 
drugs %>% 
  distinct(therapeutic_area) 

#top_levels(drugs$therapeutic_area)
## oops not labeled as a factor ... how do we change this? 
## You HAVE to do it within the line of code for top_levels --> if you do it before in a separate line of code, it will not work :/ 
top_levels(as.factor(drugs$therapeutic_area), 3) ##we can also specify if we want to show NAs or not 

```
### What is the output? 
The resulting tibble gives you the top and bottom 3 categories and the list of each item within that top and bottom 3 list. It will also tell you how many categories fall within the middle. The percentages are also calculated and can be based off of omitted NAs if that is specified.  

One downfall of this function is that factors may not be that common in data sets. One option is to use _as.factor_ function and convert as we did above.  

### Some notes to consider:  
-  you can only use top_levels() with factors! It will not work as a numeric value 
-  make sure to install the janitor package and then load the library before trying to use top_levels() function
-  the category must also have **at least** 3 levels 













